---
title: "Hansen Data Exploration"
author: "Meghan Bowen"
date: "May 18, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load_packages}

library(tidyverse) 
library(raster) ##load and manipulate rasters
library(sf) ##load shapefiles
library(sp) ##spatial objects
library(spatstat) ##work with shapefiles
library(tmap) ##map rasters + shapefiles
library(rgdal) ##for projections and coord systems

library(gfcanalysis) ##package specifically for Hansen Forestry Data!
library(landscapetools) ##for landscape plotting/viz?
library(landscapemetrics) ##R equivalent to FRAGSTATS


```


```{r load_rasters}

##load rasters

#lossyear <- raster("Hansen_GFC-2018-v1.6_lossyear_00N_080W.tif")

#plot(lossyear)


##what is projection? datum is WGS84
lossyear@crs
```


```{r load_shps}
##Muni_shp: legal Amazon boundary
##muni_2010: Brazil boundary?? not sure



##selecting "estado_id = 1" to subset
muni_shp <- read_sf(dsn = 'municipality_shapefiles', layer = "Amazon_Municipios") %>% 
  dplyr::filter(estado_id == 1) %>% 
  dplyr::select(id, estado_id)
#shp projection is also WGS84

plot(muni_shp)

##selecting pib (??) in municipios 2010
muni_2010 <- read_sf(dsn = 'municipality_shapefiles', layer = "municipios_2010") %>% 
  dplyr::select(pib)

```


```{r subset_gfc}

#using the "muni_shp" subset above (with estado_id = 1)

##calc_gfc_tiles: figure out which tiles are needed based on AOI (shp file above)
##for this function, need to convert muni_shp from multipolygon (spat out by sf) to a spatialpolygon object

#convert to spatial polygons
spd <- sf::as_Spatial(st_geometry(muni_shp),
                      IDs = as.character(1:nrow(muni_shp)))

#muni_df <- muni_shp
#muni_df$geometry <- NULL
muni_df <- as.data.frame(muni_shp)

muni_spd <- sp::SpatialPolygonsDataFrame(spd, data = muni_df)

#find the # of tiles needed. For the estado_id = 1 subset, need 4 tiles
tiles <- calc_gfc_tiles(aoi = muni_spd)
tiles

##once find the 'tiles' above, can use 'download_tiles' to grab the files from Hansen online 
##DO NOT TOUCH THIS PLZ
#dir.create("tiles_subset")

#subset of subset
output_folder <- "."
download_tiles(tiles[2], output_folder = ".", dataset = "GFC-2018-v1.6",images = c("treecover2000", "lossyear", "gain"))

##once grabbed all 'download_tiles' above, can use 'extract_gfc' to grab the "change" layers (treecover2000, loss, gain, lossyear, datamask)
##THIS IS NOT WORKING
##NEED TO WORK ON MOSAIC, make it into a RasterBrick?
gfc <- extract_gfc(aoi = muni_spd,
                   data_folder = "Malaria-Deforestation/test",
                   filename = "test.tif")

mosaic_test <- mosaic()
raster1 <- raster("tiles_subset/Hansen_GFC-2018-v1.6_datamask_00N_070W.tif")

##threshold_gfc after

##THEN can use gfc_stats

```

```{r check_subset}

##check the tiles that they have cumulative loss for each year (2001, 2002, etc)
lossyear <- raster("Hansen_GFC-2018-v1.6_lossyear_00N_070W.tif") ##FIX

treecover2000 <- raster("Hansen_GFC-2018-v1.6_treecover2000_00N_070W.tif") ##FIX

##Raster Stack: cover in 2000 and lossyear
stack <- raster::stack(treecover2000, lossyear)

###Extract by Muni_Shp
##do not runstack_ext <- extract(stack, muni_shp)

plot(lossyear)
plot(muni_shp, add = T, col ="blue")

subsetclip <- crop(lossyear, extent(muni_spd))
subset2000 <- crop(treecover2000, extent(muni_spd))

plot(subsetclip)
plot(muni_spd, add=T)

##filter by value = 1
filter_2001 <- subsetclip %in% 1

#this works?
filter2001 <- subsetclip
filter2001[filter2001 > 1] <- NA

filter_2002 <- subsetclip %in% 2

#both years
filter_0102 <- subsetclip %in% 1:2

###not working
#filter2002 <- subsetclip
#filter2002[filter2002 <= 1] <- NA

##mini extent
ext <- extent(c(-67.1, -67, -10, -9.9))
par(mfrow = c(1,2))
plot(filter_2001, ext= ext,
     main="2001",
     xlab = "lat",
     ylab = "long")
plot(filter_0102, ext=ext,
     main="2002",
     xlab = "lat",
     ylab = "long")

#test_map <- tm_shape(filter_2001) +
  tm_raster(title = "Test Map") +
  tm_shape(muni_spd) +
  tm_layout(earth.boundary = ext) +
  tm_borders(col = "darkgreen")
  
#test_map

```



```{r test_gfc}

##calc_gfc_tiles: figure out which tiles are needed based on AOI (shp file above)
##for this function, need to convert muni_shp from multipolygon (spat out by sf) to a spatialpolygon object

#convert to spatial polygons
spd <- sf::as_Spatial(st_geometry(muni_shp),
                      IDs = as.character(1:nrow(muni_shp)))

#muni_df <- muni_shp
#muni_df$geometry <- NULL
muni_df <- as.data.frame(muni_shp)

muni_spd <- sp::SpatialPolygonsDataFrame(spd, data = muni_df)

tiles <- calc_gfc_tiles(aoi = muni_spd)
tiles

##once find the 'tiles' above, can use 'download_tiles' to grab the files from Hansen online 

##once grabbed all 'download_tiles' above, can use 'extract_gfc' to grab the "change" layers (treecover2000, loss, gain, lossyear, datamask)

##threshold_gfc after

##THEN can use gfc_stats

```



```{r compress}

##expand from 0.00025 x 0.00025 by a factor of 10 --> 0.0025 X 0.0025
lossyear_res <- aggregate(lossyear, fact = 10)

##reproject?
lossyear_wgs <- projectRaster(lossyear_res, proj4string(+init=epsg:4326 +proj=longlat +ellps=WGS84))

#plot both together
plot(lossyear_res)
plot(muni_shp,add=T)

#crop muni to kinda fit this tile?
muni_bounds <- as(extent(-75, -70, -10, -4), 'SpatialPolygons')
crs(muni_bounds) <- crs(muni_2010)

crop_muni <- sf::st_crop(muni_2010, muni_bounds)

#plot both together
plot(lossyear_res)
plot(crop_muni,add=T)
```

```{r extract_mask}

##interest raster by muni?
loss_clip <- crop(lossyear_res, extent(muni_shp))

## crop and mask
lossyear_res0 <- crop(lossyear_res, extent(crop_muni))
lossyear_crop <- mask(lossyear_res0, crop_muni)

## Check that it worked
plot(lossyear_crop)
plot(crop_muni, add=TRUE, lwd=2)

```



```{r tmap}

##test map 1: border of shapefile ontop of raster
test_map <- tm_shape(lossyear_res) +
  tm_raster(title = "Test Map") +
  tm_shape(muni_shp) +
  tm_borders(col = "darkgreen")
  
test_map

##test map 2: "extract by mask" (lossyear_crop)
##not working?
test_map2 <- tm_shape(lossyear_crop) +
  tm_raster(title = "Test Map") +
  tm_shape(muni_shp) +
  tm_borders(col = "darkgreen")
  
test_map2

```


```{r explore_landscapemetrics}

##show landscape with all patches/etc
show_landscape(lossyear_crop)

```


##General Spatial R Questions:

1. **How to get these huge tile rasters down?! even by fact = 10 it took 20 minutes to get it down. Do we even want to get it down?**
2. can't use tm_shape with a shapefile with multiple columns? need to filter it to one attribute?
3. need a landscape raster? ie with patches, habitat types ...? or will the lossyear be the "landscape" and the patches/areas are in %loss of forest???

- double check the projection (make sure it's ESPG: 4326)
- nine states in the legal amazon (wikipedia)

## Landscape Metrics Notes

- Area and edge metrics describe the size of patches and classes and the amount of edge. An edge is defined as the border between two patches of class i and k in meters. These metrics mainly characterize the composition of the landscape and are able to show dominance or rareness of classes.
- Shape metrics describe the shape of patches, mainly by using its area and perimeter. This can be important for many research questions, because e.g. even though, being equal in size, long and narrow patches has probably different characteristics than a squared patch of the same size.
- Core metrics describe the area of patches that are not an edge. These metrics can be interesting for research questions in which e.g. only areas that are not influenced by neighboring patches of a different class are of interest.
- Aggregation metrics describe if patches (of the same class) are rather clumped (aggregated) or tend to be isolated. Following, these metrics describe mainly the spatial configuration of the landscape.


## Notes from June 4
- patch level info
- how to get the patch level calculations
- raster based on one outline of one municipality
- here's all of the raster data, here's all the municipality, run each municipality for each year (2001, 2001-2002, etc...)
- 30x30 m is fine 
